---
title: "Plotting with ggplot"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: yes
    allow_skip: false
    toc: true
    toc_depth: 2
    toc_float: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
# first, some packages for interactive tutorial
library(knitr)
library(learnr)
library(gradethis) # not really necessary if we're not interested in automatic code grading (and answers vary wildly!)
library(ggplot2)
# function to record events
tutorial_event_recorder <- function(tutorial_id, tutorial_version, user_id, 
                                    event, data) {
  cat(tutorial_id, " (", tutorial_version, "): ", user_id , "\n", sep = "")
  cat("event: ", event, "\n", sep = "")
  if (!is.null(data[['code']])) {
    time_stamp <- format(Sys.time(), "%a %b %d %X %Y")
    print(paste(time_stamp, data$code, data$error_message))
    # store timestamp, code, and any errors in a file 
    # TODO change format to something better like csv
    fileConn <- file("executions.txt", open="at")
    writeLines(paste("----\n", time_stamp, "\n----\n", data$code, "~~~~\nError: ", 
                    ifelse(!is.null(data$error_message), data$error_message, "none")), fileConn)
    close(fileConn)
  }
}
options(tutorial.event_recorder = tutorial_event_recorder)
tutorial_options(exercise.timelimit = 60, exercise.completion = TRUE, exercise.checker = gradethis::grade_learnr)
# by default, let's not show code as many are written for learnr
opts_chunk$set(comment = NA)
```

```{r echo=FALSE}
# declare any dataframes here
mpg <- ggplot2::mpg
```

## Introduction

In this experiment, you will recreate two different plots using R's `ggplot` library. Before that, let's do a quick introduction on how to plot using `ggplot`.

To make a graph in `ggplot` this is the basic template:

```r
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

where `<DATA>` is the dataframe, `<GEOM_FUNCTION>` is the type of plot and `<MAPPINGS>` are the x and y variables. To make a graph, you need to replace the bracketed sections in the code below with a dataset, a geom function, or a collection of mappings.

Let's look at an example with the **mpg** dataset. This dataset is fuel economy data from 1999 and 2008 for 38 popular models of car. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car. Here it is:

```{r}
mpg
```

Here is an example of a scatterplot of **mpg** data, where we look at the relationship between `displ` (as our x) and `hwy` (as our y):

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Going back to our template, note the following:

- The `<DATA>` is replaced with **mpg**
- The `<GEOM_FUNCTION>` is replaced with `geom_point` for scatterplot*
- The `<MAPPINGS>` is replaced with `x = displ, y = hwy`, our x and y variables

*Keep in mind that you can use multiple geom functions.

Click on the 'Next Topic' button to move on to your Task 1.

## Task 1

Using the same data **mpg** as before:

```{r echo=FALSE}
mpg
```


<!--Source: https://ggplot2.tidyverse.org/reference/geom_smooth.html-->
Recreate the following plot:

```{r echo=FALSE}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```


```{r task1, exercise = TRUE, exercise.lines = 5}



```

Once you are done or if you would like to move on, click on the 'Next Topic' button for Task 2.

## Task 2

Using the same data **mpg** as before:

```{r echo=FALSE}
mpg
```

<!--Source: https://ggplot2.tidyverse.org/reference/geom_bar.html-->
Recreate the following plot:

```{r echo=FALSE}
ggplot(mpg, aes(class)) + 
  geom_bar(aes(fill = drv), position = position_stack(reverse = TRUE)) +
  coord_flip() +
  theme(legend.position = "top")
```

```{r task2, exercise = TRUE, exercise.lines = 5}

```


