---
title: "Plotting with ggplot"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: no
    allow_skip: false
    toc: false
    toc_float:
      collapsed: false
      smooth_scroll: false
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
# first, some packages for interactive tutorial
library(knitr)
library(learnr)
library(gradethis) # not really necessary if we're not interested in automatic code grading (and answers vary wildly!)
library(ggplot2)
library(rvest)
# function to record code execution events
tutorial_event_recorder <- function(tutorial_id, tutorial_version, user_id, event, data) {
  if (!is.null(data[['code']]) && 
      all(unlist(lapply(strsplit(data[['code']], "\n"), function(x) ifelse(x == "", FALSE, TRUE)))) && 
      data[['code']][[1]] != "") {
    time_stamp <- format(Sys.time(), "%a %b %d %X %Y")
    print(paste(time_stamp, data$code, data$error_message))
    # store in csv file: time, code, error, warning
    print(as.character(time_stamp))
    x <- data.frame(
          time = as.character(time_stamp), 
          code = as.character(data$code), 
          error = ifelse(!is.null(data$error_message), data$error_message, ""),
          warning = ""
        )
    # if there is no error message, check if there is a warning
    if (is.null(data$error_message)) {
      html_file <- data$output
      page <- read_html(html_file)
      pre_node <- html_node(page, css='pre')  
      # if there's warning store the message, else leave it blank
      x[['warning']] = ifelse(grepl("Warning", pre_node, fixed=TRUE), html_text(pre_node), "")
    }
    write.table(x, file = "PX_Executions.csv", sep = ",", col.names = !file.exists("PX_Executions.csv"), row.names = FALSE, append=TRUE)
  }
}
options(tutorial.event_recorder = tutorial_event_recorder)
tutorial_options(exercise.timelimit = 60, exercise.completion = TRUE)
# by default, let's not show code as many are written for learnr
opts_chunk$set(comment = NA, fig.width=4, fig.height=3)
```

```{r echo=FALSE}
# declare any dataframes here
mpg <- ggplot2::mpg
```

## Warm up

In this experiment, you will recreate three different plots using R's `ggplot` library. Before that, let's do a quick intro on using `ggplot` and the code editor in this. To make a graph in `ggplot` this is the basic template:

```r
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

where `<DATA>` is the dataframe, `<GEOM_FUNCTION>` is the type of plot and `<MAPPINGS>` are the x and y variables. Let's look at an example with the **mpg** dataset. This dataset is fuel economy data from 1999 and 2008 for 38 popular models of car:

```{r echo=FALSE}
head(mpg, 3)
```

Here is an example of a scatterplot of **mpg** data, where we look at the relationship between `displ` (x) and `hwy` (y). 
Execute the following by clicking on "Run Code" (if you ever want to reset code, click "Start over"):

```{r task0, exercise = TRUE, exercise.lines = 2, fig.width=4, fig.height=2}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Going back to our template, note the following:

- The `<DATA>` is replaced with **mpg**
- The `<GEOM_FUNCTION>` is replaced with `geom_point` for scatterplot*
- The `<MAPPINGS>` is replaced with `x = displ, y = hwy`, our x and y variables

*Keep in mind that you can use multiple geom functions.

Click on the 'Next Topic' button to move on to your Task 1.

## Task 1: Data

Using the same data **mpg** as before:

```{r echo=FALSE}
head(mpg, 5)
```

## Task 1: Plot
<!--Based on: https://ggplot2-book.org/getting-started.html#basic-use-->
Recreate the following plot:

```{r echo=FALSE, fig.width=4, fig.height=3}
ggplot(mpg, aes(drv, cty)) + 
  geom_violin()
```


```{r task1, exercise = TRUE, exercise.lines = 5, fig.width=4, fig.height=3}



```


<!-- ```{r task1-check} -->
<!-- grade_result( -->
<!--   pass_if(~ .result != "", "Great! Please move on to the next task."), -->
<!--   fail_if(~ .result == "", "No code was submitted.") -->
<!-- ) -->
<!-- ``` -->

Once you are done or if you would like to move on, notify the experimenter you are ready for Task 2.

<!-- TODO: separate out task 2 and task 3 into their own Rmds -->

<!-- ## Task 2 -->

<!-- Using the same data **mpg** as before: -->

<!-- ```{r echo=FALSE} -->
<!-- mpg -->
<!-- ``` -->

<!--Based on: https://ggplot2-book.org/getting-started.html#axes-->
<!-- Recreate the following plot: -->

<!-- ```{r echo=FALSE} -->
<!-- ggplot(mpg, aes(drv, hwy)) + -->
<!--   geom_jitter(width = 0.25, alpha = 1 / 2) + -->
<!--   xlab("Wheel drive") +  -->
<!--   ylab("Highway mpg") -->
<!-- ``` -->

<!-- ```{r task2, exercise = TRUE, exercise.lines = 5} -->

<!-- ``` -->

<!-- Once you are done or if you would like to move on, click on the 'Next Topic' button for Task 3. -->

<!-- ## Task 3 -->

<!-- Using the same data **mpg** as before: -->

<!-- ```{r echo=FALSE} -->
<!-- mpg -->
<!-- ``` -->

<!--
Based on
https://ggplot2.tidyverse.org/reference/geom_bar.html https://ggplot2-book.org/getting-started.html#qplot-facetting
-->
<!-- Recreate the following plot: -->

<!-- ```{r echo=FALSE} -->
<!-- ggplot(mpg, aes(class)) +  -->
<!--   geom_bar(aes(fill = drv), position = position_stack(reverse = TRUE)) + -->
<!--   coord_flip() + -->
<!--   facet_wrap(~cyl) + -->
<!--   theme(legend.position = "top") -->
<!-- ``` -->

<!-- ```{r task3, exercise = TRUE, exercise.lines = 5} -->

<!-- ``` -->


